% IEEE Paper Template for US-LETTER Page Size (V1)
% Sample Conference Paper using IEEE LaTeX style file for US-LETTER pagesize.
% Copyright (C) 2006-2008 Causal Productions Pty Ltd.
% Permission is granted to distribute and revise this file provided that
% this header remains intact.
%
% REVISION HISTORY
% 20080211 changed some space characters in the title-author block
%
% \documentclass[10pt,conference,letterpaper]{IEEEtran}
% \usepackage{times,amsmath,epsfig}
% \usepackage[utf8]{inputenc}
% \usepackage[usenames,dvipsnames]{color}
% \usepackage{amssymb}

\documentclass[a4paper,english]{llncs}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{lmodern}
\usepackage{float}
\usepackage{slashed}
\usepackage{graphicx}
\usepackage{fixltx2e}
\usepackage{amsmath}
\usepackage{epstopdf}
\usepackage{url}
\usepackage{seqsplit} % for splitting very long variable names
\usepackage[dvipsnames]{xcolor}
\usepackage{listings}

\hyphenation{bio-ban-kers}
\hyphenation{Bio-bank-Cloud}

\newcommand\YAMLcolonstyle{\color{black}}
\newcommand\YAMLkeystyle{\color{blue}}
\newcommand\YAMLvaluestyle{\color{black}}

% \newcommand\language@yaml{yaml}
% 
% \expandafter\expandafter\expandafter\lstdefinelanguage
% \expandafter{\language@yaml}
% {
%   keywords={name,ec2,type,region,cookbooks,groups,size,recipes,github,branch},
%   keywordstyle=\color{blue}\bfseries,
%   basicstyle=\color{black},                                 % assuming a key comes first
%   sensitive=false,
%   comment=[l]{\#},
%   morecomment=[s]{/*}{*/},
%   commentstyle=\color{purple}\ttfamily,
%   stringstyle=\YAMLvaluestyle\ttfamily,
%   moredelim=[l][\color{orange}]{\&},
%   moredelim=[l][\color{magenta}]{*},   % switch to value style at :
%   literate =    {---}{{\ProcessThreeDashes}}3
%                 {>}{{\textcolor{red}\textgreater}}1     
%                 {|}{{\textcolor{red}\textbar}}1 
%                 {\ -\ }{{\mdseries\ -\ }}3,
% }
% 
% % switch to key style at EOL
% \lst@AddToHook{EveryLine}{\ifx\lst@language\language@yaml\YAMLkeystyle\fi}
% % \makeatother
% 
% \newcommand\ProcessThreeDashes{\llap{\color{cyan}\mdseries-{-}-}}
% 
% \usepackage{hyperref}
% \newcommand\fnurl[2]{%
% \footnote{\label{#1}\href{#2}{#2}}%
% }
% 



\makeatletter
\raggedbottom % remove unwanted space between paragraphs 

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
\pdfpageheight\paperheight
\pdfpagewidth\paperwidth

\floatstyle{ruled}
\newfloat{algorithm}{tbp}{loa}
\providecommand{\algorithmname}{Algorithm}
\floatname{algorithm}{\protect\algorithmname}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\usepackage{paralist}
\usepackage[noend]{algorithm,algcompatible,algpseudocode}
\usepackage{amssymb} % for symbols 
\usepackage{caption} % algorthm is a float and does not split. use caption instead
% \usepackage{subcaption}
\usepackage{etoolbox}\AtBeginEnvironment{algorithmic}{\footnotesize}
%\algsetup{linenosize=\small}
\usepackage{subfig}
\usepackage[T1]{fontenc}
\usepackage{mwe}    % loads »blindtext« and »graphicx«

\makeatother
\algblockdefx[NAME]{StartTransaction}{EndTransaction}
[1]{\textbf{begin transaction} \textbf{#1}}{\textbf{commit transaction}}
\algrenewcommand\alglinenumber[1]{\footnotesize #1:}

\usepackage{babel}
\usepackage{cite}
\usepackage{amssymb}% http://ctan.org/pkg/amssymb
\usepackage{pifont}% http://ctan.org/pkg/pifont
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{53}}%

\title{BiobankCloud: a Platform for the Secure Storage, Sharing, and Processing of Large Biomedical Data Sets}

\author{Alysson Bessani\inst{5}, J\"{o}rgen Brandt\inst{2}, Marc Bux\inst{2}, Vinicius Cogo\inst{5}, Lora Dimitrova\inst{4}, Jim Dowling\inst{1}, Ali Gholami\inst{1}, Kamal Hakimzadeh\inst{1}, Micheal Hummel\inst{4}, Mahmoud Ismail\inst{1}, Erwin Laure\inst{1}, Ulf Leser\inst{2}, Jan-Eric Litton\inst{3}, Roxanna Martinez\inst{3}, Salman Niazi\inst{1}, Jane Reichel\inst{6}, Karin Zimmermann\inst{4}}

\institute{KTH - Royal Institute of Technology,\\
\email{\{jdowling, gholami, mahh, maism, erwinl, smkniazi\}@kth.se}
\and
Humboldt University\\
\email{\{leser, bux, joergen.brandt\}@informatik.hu-berlin.de}
\and
Karolinska Institute\\
\email{\{Jan-Eric.Litton, Roxanna.Martinez\}@ki.se}
\and
Charite\\
\email{\{Michael.Hummel, Lora.Dimitrova, Karin.Zimmermann\}@charite.de}
\and
LaSIGE, Faculdade de Ci\^{e}ncias, Universidade de Lisboa, Portugal\\
\email{\{bessani, vielmo\}@lasige.di.fc.ul.pt}
\and
Uppsala University\\
\email{\{jane.reichel\}@jur.uu.se}
}

\newif\ifshowcomments
\showcommentstrue

\ifshowcomments
\newcommand{\mynote}[2]{\fbox{\bfseries\sffamily\scriptsize{#1}}
{\small$\blacktriangleright$\textsf{\emph{#2}}$\blacktriangleleft$}}
\else
\newcommand{\mynote}[2]{}
\fi
\newcommand{\jim}[1]{\textcolor{Red}{\mynote{Jim}{#1}}}

\begin{document}
\maketitle

\begin{abstract}
Biobanks store and catalog human biological material that is increasingly being digitized using next-generation sequencing (NGS). There is, however, a computational bottleneck, as existing software systems are not scalable and secure enough to store and process the incoming wave of genomic data from NGS machines. In the BiobankCloud project, we are building a Hadoop-based platform for the secure storage, sharing, and parallel processing of genomic data. We extended Hadoop to include support for multi-tenant studies, reduced storage requirements with erasure coding, and added support for extensible and consistent metadata. On top of Hadoop, we built a scalable scientific workflow engine featuring a proper workflow definition language focusing and simple integration and chaining of existing tools, adaptive scheduling on Apache Yarn, and support for iterative dataflows. Our platform also supports the secure sharing of data across different, distributed Hadoop clusters. The software is easily installed and comes with a provide user-friendly web interface for running, managing, and accessing data sets behind a secure 2-factor authentication. Initial tests have shown that the engine scales well to dozens of nodes. The entire system is open-source and includes pre-defined workflows for popular tasks in biomedical data analysis, such as variant identification, differential transcriptome analysis using RNA-Seq, and analysis of miRNA-Seq, and ChIP-Seq data.
\end{abstract}

%\vskip-5pt
\input{introduction}
%\vskip-5pt
\input{relatedWork}
% \input{regulation}
%\vskip-5pt
\input{lims}
%\vskip-5pt
\input{security}
%\vskip-5pt
\input{hops}
%\vskip-5pt
\input{sassfee}
%\vskip-5pt
\input{workflows}
%\vskip-5pt
\input{charonfs}
% \vskip-5pt
\input{karamel}

\section{Conclusions}
In this paper, we introduced the BiobankCloud platform, which provides many features necessary for biobanks to adopt Hadoop-based solutions for managing NGS data. Critical security features that we have introduced for managing sensitive data include multi-tenancy to isolate Studies and 2-factor authentication. Next-generation data management systems for NGS data must be massively scalable. We introduced our scalable storage service, HopsFS, our processing framework, HopsYARN, and our framework for scalable bioinformatics workflows, SAASFEE. We also provide metadata design and search services, while ensuring the integrity of metadata. Finally, in Charon, we showed how we can leverage public clouds to share data securely between clusters. BiobankCloud's secure and scalable open platform software is helping to remove the biobank bottleneck and enabling population-level genomics.
% \vskip-5pt
\section{Acknowledgements}
This work is funded by the EU FP7 project ``Scalable, Secure Storage and Analysis of Biobank Data'' under Grant Agreement no. 317871. 
%\vskip-5pt
\bibliographystyle{ieeetr}
\bibliography{bbc-dmah15}

\end{document}
