\section{Related work}

Our platform covers a broad set of technologies providing components that solve problems in the areas of security in Hadoop, sharing data, parallel data processing, and data management. Here, we focus on related work in the area of parallel data analytics and discuss some general and some domain-specific solutions. I general, research in platforms for large-scale data analysis is flourishing over the last years. Application-independent systems like Spark~\cite{Zaharia2010} or Stratosphere~\cite{Alexandrov2014} focus on supporting efficient and distributed data analysis, but focus on query evaluation, especially for relational queries. This leads to some design choices that make supporting scientific analysis pipelines (or scientific workflows) rather cumbersome. % ???REF
In contrast, specialized scientific workflow management systems like Taverna, Kepler, or Galaxy, typically neglect the Big Data aspect, i.e., are not able to scale workflow executions efficiently over large clusters~\cite{Bux2013}. 

The lack of a flexible and open platform for running complex analysis pipelines over large data sets let to the development of a number of highly specialized systems. For instance, tools like Crossbow~\cite{Langmead2009b} perform one specific operation (e.g. sequence alignment) in a distributed manner, but cannot be used for any other type of analysis. Also for the most important types of complex analysis, like sequence assembly or mutation detection and evaluation, specialized systems exist. Some very popular systems, like GATK~\cite{McKenna2010}, only supported a very simple form of parallelism and cannot fully exploit existing computational resources. Technically more advanced solutions like Adam~\cite{adam15}, Halvade~\cite{halvade15}, Seal~\cite{seal11}, and PigSeq~\cite{seqpig14} show much better scaling, but are not general purpose solutions but specifically developed for one type of analysis. Appropriate models of data security and privacy are not supported in any of the systems we are aware of as they all delegate these issues to the file system level, which is especially problematic of HDFS is used, although frameworks to enhance Hadoop with access control do exist; for instance, Apache Ranger provides an attribute-based access control server that is general and expressive enough to be applied to most Hadoop services but not fast enough to authorize HDFS operations at the NameNode or application requests at the ResourceManager. Overall, we see a clear lack of a flexible, open, secure, and scalable platform for scientific data analysis on large data sets.
